
```{r}
library (Seurat)
library(dplyr)
library(MetBrewer)
library(presto)
library(gplots)
library(RColorBrewer)
library(ggplot2)
library(simspec)
library(SeuratWrappers)
library(tidyr)
library(dplyr)

```

                                    #### pre-processing steps

       
          #Seurat <- NormalizeData (Seurat, normalization.method = "LogNormalize", scale.factor = 10000)

          # Select the top 1000 most variable features
          #Seurat <- FindVariableFeatures (Seurat, nfeatures = 3000, selection.method = "vst")

          #Seurat <- subset (Seurat, subset = nFeature_RNA > 500 & percent.mt < 10)
          #Seurat <- ScaleData (Seurat)

          #Seurat <- RunPCA (Seurat, features = VariableFeatures (Seurat)) 

          #Seurat <- FindNeighbors (Seurat, reduction = "pca", dims = 1:20)
          #Seurat <- FindClusters (Seurat, resolution = 0.8)

                          ########## HTO classification #######
                          
                          Condition == "1" ~ "control_Mat",
                          Condition == "2" ~ "WLSKO_Mat",
                          Condition == "3" ~ "control_NoMat",
                          Condition == "4" ~ "WLSKO_NoMat",
                          Condition == "5" ~ "WLSKO_Chir_Mat",
                          Condition == "6" ~ "WLSKO_Py60_Mat",
                          Condition == "7" ~ "control_Chir_Mat",
                          Condition == "8" ~ "control_DMSO_Mat",
                          Condition == "9" ~ "WLSKO_DMSO_Mat",
                          Condition == "10" ~ "control_Py60_Mat"

```{r}
seurat_obj_annotated <- readRDS( "seurat_obj_processed_WLS_KO.rds")

```


                                                   ##### UMAP

```{r}
Idents(seurat_obj_annotated) <- "seurat_clusters"

seurat_obj_annotated<-RunUMAP(seurat_obj_annotated, reduction = "pca", dims = 1:22, min.dist =0.4, spread = 0.6)

plot1 <- UMAPPlot(seurat_obj_annotated, group.by = "Treatment", cols =met.brewer("Cross", n = 10) ,  pt.size = 0.5, order= T) & NoAxes()
plot1

plot2 <- UMAPPlot(seurat_obj_annotated, group.by = "dataset", cols =met.brewer("Johnson", n = 2) ,  pt.size = 0.3, order= T) & NoAxes()
plot2

plot3 <- UMAPPlot(seurat_obj_annotated, group.by = "Cell_line", cols = c( "#282626","#BFBFBF") ,  pt.size = 0.3, order= T) & NoAxes()
plot3

```

                                  ## Annotate clusters

```{r}

Idents(seurat_obj_annotated) <- "seurat_clusters"

# stash current cluster IDs
seurat_obj_annotated[["cluster.ids"]] <- Idents(object = seurat_obj_annotated)

# enumerate current cluster IDs and the labels for them
cluster.ids <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11,12,13,14,15,16,17,18)

anno_class1 <- c("Telencephalic neurons", #  0
            "Telencephalic progenitors",#  1
            "Choroid Plexus", #F   2
            "Telencephalic neurons", #    3
            "Telencephalic progenitors",#   4
            "Telencephalic neurons", # ventral    5
            "Mid-hind neurons", #    6
            "Telencephalic neurons",#    7
            "Telencephalic progenitors",#ventral    8
            "Telencephalic progenitors",#   Dorsal 9
            "Telencephalic progenitors", #   10
            "Choroid Plexus", #    11
            "Retinal progenitors", #       12
            "Mid-hind progenitors", #  13
            "removed",        #   14
            "Neural crest", #    ventral 15
            "Mid-hind progenitors",   #    16
            "Mid-hind progenitors", #17
            "Choroid Plexus") # 18
 

seurat_obj_annotated@meta.data[,'anno_class1'] <- plyr::mapvalues(x = seurat_obj_annotated@meta.data$cluster.ids, from = cluster.ids, to = anno_class1)

seurat_obj_annotated$anno_class1 <- factor(seurat_obj_annotated$anno_class1, levels = c("Telencephalic progenitors","Telencephalic neurons","Retinal progenitors","Choroid Plexus","Mid-hind progenitors","Mid-hind neurons","Neural crest"))

#color_anno = met.brewer("Cassatt2",9)
color_anno = c("#049983","#054","#264653","#277891","#930770","#6B081F","#f4a261","#cc9933","#277781")


classdimplot <- DimPlot(seurat_obj_annotated, reduction = "umap", pt.size = 0.5, label = F,order = F, group.by = 'anno_class1', cols = color_anno) & NoAxes()

classdimplot

```


```{r}

FeaturePlot_3 <-  FeaturePlot(seurat_obj_annotated, c( "WLS","LHX9","FOXG1","TTR"), ncol=2, pt.size = 1, cols = colorRampPalette(c("lightgrey","turquoise4","midnightblue"))(30), order=T)& NoAxes() &NoLegend()
FeaturePlot_3

```


```{r}
library(dittoSeq)
barplot_conditions<-dittoBarPlot(
  object = seurat_obj_annotated,
  var = "anno_class1",
  group.by = "Treatment",
  color.panel = color_anno,
  retain.factor.levels = T,
  ylab = "Cluster proportion")
barplot_conditions
```                 
                        
                     
```{r}
#####
dotplot <- function(exprmat,
                    genes,
                    cl,
                    gene.reorder=TRUE,
                    specified.order=NULL,
                    cl.reorder.by.hc=T,
                    max.diag=TRUE,
                    emphasize_genes = NULL,
                    emphasize_cl = NULL,
                    colors=c("#d9d9d9", "#252525"),
                    point.size.factor=3.5,
                    plot.margin=NULL,
                    plot.cex=0.5,
                    col.space.factor=1,
                    row.space.factor=1,
                    verbose = F)



##################################
Idents(seurat_obj_annotated) <- "anno_class1"


genes_2 <- c("SOX2","FOXG1","NKX2-1","DLX2","MAP2","LHX9","IRX3","HOXB2","SOX10","FOXD3","TWIST1","DCN","RAX","SIX6","PAX6","TTR","WLS","OTX2","LMX1A","ZIC1","MSX1")

dotplot_1 <- dotplot(seurat_obj_annotated@assays$RNA@data[genes_2,], genes_2, seurat_obj_annotated$seurat_clusters)

dotplot_1 <- DotPlot(object = seurat_obj_annotated, features = genes_2, cluster.idents = T, dot.scale = 6,dot.min = 0,scale.by = "radius", col.min = 0) + theme_classic()+ theme(axis.text.x = element_text(angle = 90))+scale_colour_gradient2(low="grey1", mid="grey100", high="midnightblue")+
  geom_point(aes(size=pct.exp), shape = 21, colour="black", stroke=0.2)
dotplot_1

```
                     
                        ######### differential abundance #####
                        
```{r}
 freq <- table(seurat_obj_annotated$anno_class1, seurat_obj_annotated$Treatment)
prop <- apply(freq,2,function(x) x/sum(x))

layout(matrix(1:3,nrow=1)); par(mar=c(8,5,1,1))
barplot(freq, col=color_anno,
        border=NA, las=2, ylab="Frequency", cex.names = 0.8)
barplot(prop, col=color_anno,
        border=NA, las=2, ylab="Proportion", cex.names = 0.8)
plot.new()

legend("left", fill=color_anno, legend=rownames(freq), bty="n")       

```

```{r}
freq_fisher <- function(Treatment, have_identity){
  freq <- table(factor(have_identity, levels=c(TRUE,FALSE)),
                Treatment)
  test <- fisher.test(freq)
  res <- setNames(c(test$estimate, test$p.value), c("oddsratio","pval_fisher"))
  return(res)
}

anno_class1_enrichment <- data.frame(anno_class1 = levels(seurat_obj_annotated$anno_class1),
  # loop and test every anno_class1
  t(sapply(levels(seurat_obj_annotated$anno_class1), function(anno_class1)
    freq_fisher(Treatment = factor(seurat_obj_annotated$Treatment, levels=c("control_Py60_Mat","WLSKO_Py60_Mat")),
                have_identity = seurat_obj_annotated$anno_class1 == anno_class1)
  )),
  row.names=NULL)
anno_class1_enrichment$padj_fisher <- p.adjust(anno_class1_enrichment$pval_fisher)

anno_class1_enrichment


```


```{r}
#############################       Volcano plot DE       ##############################
Idents(seurat_obj_annotated) <- "Condition"

cl_markers_control.WLS_KO_Chir <- FindMarkers(seurat_obj_annotated, ident.1 =c("7"), ident.2 = c("5"), only.pos = F)

cl_markers_control.WLS_KO_Py60<- FindMarkers(seurat_obj_annotated, ident.1 =c("10"), ident.2 = c("6"), only.pos = F)


library(EnhancedVolcano)

volcano_1 <- EnhancedVolcano(cl_markers_control.WLS_KO_Chir,
                             lab = rownames(cl_markers_control.WLS_KO_Chir), 
                             #selectLab = rownames(cl_markers_Ctrl.Py60),
                           #selectLab =
                            # c("WLS","FOXG1","EMX2","SIX3","SIX6","WNT4","WNT2B","FGF17","PAX3","PAX7","OTX2","EN2","IRX3","IRX2","CYP26A1","TFAP2A","HOXB2","GBX2","HAPLN1","PRTG","FOXH1","HESX1","SOX21","SFRP2","SOX10","FOXD3","POU5F1","SOX2","GPC3","COL4A1","NRG1","LHX2","HAPLN3","FAT1","IGFBP2","FGF8","FEZF1","FEZF2","DLK1","FZD5","ZIC2","PRTG","BMP4","BMP7","HYAL2","POU4F1","RSPO3","WNT5A","FJX1","FZD8"),
                        c("WLS","FOXG1","SIX3","WNT4","WNT2B","FGF17","PAX3","PAX7","OTX2","EN2","IRX3","IRX2","CYP26A1","TFAP2A","HOXB2","GBX2","HAPLN1","PRTG","FOXH1","HESX1","SOX21","SFRP2","SOX10","FOXD3","POU5F1","SOX2","SOX4","GPC3","COL4A1","NRG1","LHX2","HAPLN3","FAT1","IGFBP2","FGF8","FEZF1","FEZF2","DLK1","FZD5","ZIC2","PRTG","BMP4","BMP7","HYAL2","POU4F1","RSPO3","RSPO2","PCP4","GJA1","WNT5A", "LMX1B","LMX1A","TTR","SHH","EN1","WNT5A","MSX2","CCN1","SOX4","MARCKS","STMN2"),
                             x = 'avg_log2FC',
                             y = 'p_val',
                             title = 'control WLS_KO Chir',
                             pCutoff = 10e-3,
                             pCutoffCol = "p_val_adj",
                             FCcutoff = 0.5,
                             xlim = c(-3, 3),
                             pointSize = 3,
                             col=c("maroon3","purple4","grey60","black"),
                             drawConnectors = T)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+ theme(text = element_text(size = 8))    

volcano_1

ggsave( "volcano_cl_markers_control_WLS_KO_Chir.pdf",volcano_1,width=40, height=25, units="cm", dpi = 300 )


```

               

  
